Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit

### 2024-06-20 optim.qmd
pseudo algo update

To solve the optimisation problem embedded in the PP guided tour, a starting projection, an index function and the maximum number of iterations are provided as input. Then, the current projection is evaluated by the index function. The projection is then moved in a direction determined by a random factor, influenced by how far along we are in the optimization process. Occasionally, we might explore completely new directions like a jellyfish might with ocean currents. A new projection is accepted if it is an improvement compared to the current one, rejected otherwise.
This process continues and iteratively improves the projection, until the pre-specified maximum number of trials is reached.

@@

**Input**: `current_projections`, `index_function`, `tries`, `max_tries`

@@

**Initialize** `best_jelly` as the projection with the best index value
from `current_projections`, and `current_index` as the array of index
values for each projection in `current_projections`

**for** each try in 1 to max_tries **do**

> Calculate $c_t$ based on the current try and max_tries

> **if** $c_t$ is greater than or equal to $0.5$ **then**
>
> > Define trend based on the best jelly and current projections
>
> > Update each projection towards the trend using a random factor and
> > orthonormalisation
>
> **else**
>
> > **if** a random number is greater than $1 - c_t$ **then**
> >
> > > Slightly adjust each projection with a small random factor (Type A
> > > passive)
> >
> > **else**
> >
> > > For each projection, compare with a random jelly and adjust
> > > towards or away from it (Type B active)
>
> Update the orientation of each projection to maintain consistency
>
> Evaluate the new projections using the index function

> **if** any new projection is worse than the current, revert to the
> `current_projection` for that case
>
> > Determine the projection with the best index value as the new
> > best_jelly

> **if** the try is the last one, print the final best projection and
> **exit**

**return** the set of projections with the updated best_jelly as the
optimized_projection
:::

@@

The JSO implementation involves several key parameters that control its
search process in optimization problems. These parameters are designed
to guide the exploration and exploitation phases of the algorithm. While
the specific implementation details can vary depending on the version of
the algorithm or its application, we focus on two main parameters that
are most relevant to our application: the number of jellyfish and drift.

>>>>

To solve the optimisation problem embedded in the PP guided tour, a starting projection, an index function and the maximum number of iterations are provided as input. Then, the current projection is evaluated by the index function. The projection is then moved in a direction determined by a random factor, influenced by how far along we are in the optimization process. Occasionally, completely new directions may be taken like a jellyfish might with ocean currents. A new projection is accepted if it is an improvement compared to the current one, rejected otherwise.
This process continues and iteratively improves the projection, until the pre-specified maximum number of trials is reached.

@@

**Input**: `current_projections`, `index_function`, `trial_id`, `max_trial`

@@

**Initialize** `current_best` as the projection with the best index value
from `current_projections`, and `current_idx` as the array of index function
values for each projection in `current_projections`

**for** each trial_id in 1 to max_tries **do**

> Calculate $c_t$ based on the current_idx and max_trial

> **if** $c_t$ is greater than or equal to $0.5$ **then**
>
> > Define trend based on the current_best and current_projections
>
> > Update each projection towards the trend using a random factor and
> > orthonormalisation
>
> **else**
>
> > **if** a random number is greater than $1 - c_t$ **then**
> >
> > > Slightly adjust each projection with a small random factor (passive)
> >
> > **else**
> >
> > > For each projection, compare with a random jelly and adjust
> > > towards or away from it (active)
>
> Update the orientation of each projection to maintain consistency
>
> Evaluate the new projections using the index function

> **if** any new projection is worse than the current, revert to the
> `current_projections` for that case
>
> > Determine the projection with the best index value as the new
> > current_best

> **if** trial_id $\geq$ max_trial, print the last best projection 
> **exit**

**return** the set of projections with the updated current_best as the
optimized_projection
:::

@@

The JSO implementation involves several key parameters that control its
search process in optimization problems. These parameters are designed
to guide the exploration and exploitation phases of the algorithm. While
the specific implementation details can vary depending on the version of
the algorithm or its application, we focus on two main parameters that
are most relevant to our application: the number of jellyfish and the maximum number of trials.

### 2024-06-20 optim.tex
pseudo algo update

\begin{keyword}
    projection pursuit \sep Jellyfish Search Optimiser (JSO) \sep 
    Optimisation
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

To solve the optimisation problem embedded in the PP guided tour, a
starting projection, an index function and the maximum number of
iterations are provided as input. Then, the current projection is
evaluated by the index function. The projection is then moved in a
direction determined by a random factor, influenced by how far along we
are in the optimization process. Occasionally, we might explore
completely new directions like a jellyfish might with ocean currents. A
new projection is accepted if it is an improvement compared to the
current one, rejected otherwise. This process continues and iteratively
improves the projection, until the pre-specified maximum number of
trials is reached.

\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, opacitybacktitle=0.6, colback=white, left=2mm, coltitle=black, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, leftrule=.75mm, bottomtitle=1mm, arc=.35mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, opacityback=0]

\textbf{Input}: \texttt{current\_projections}, \texttt{index\_function},
\texttt{tries}, \texttt{max\_tries}

@@

\textbf{Initialize} \texttt{best\_jelly} as the projection with the best
index value from \texttt{current\_projections}, and
\texttt{current\_index} as the array of index values for each projection
in \texttt{current\_projections}

\textbf{for} each try in 1 to max\_tries \textbf{do}

\begin{quote}
Calculate \(c_t\) based on the current try and max\_tries
\end{quote}

@@

\begin{quote}
Define trend based on the best jelly and current projections
\end{quote}

@@

\begin{quote}
Slightly adjust each projection with a small random factor (Type A
passive)
\end{quote}

@@

\begin{quote}
For each projection, compare with a random jelly and adjust towards or
away from it (Type B active)
\end{quote}
\end{quote}

@@

\begin{quote}
\textbf{if} any new projection is worse than the current, revert to the
\texttt{current\_projection} for that case

\begin{quote}
Determine the projection with the best index value as the new
best\_jelly
\end{quote}
\end{quote}

\begin{quote}
\textbf{if} the try is the last one, print the final best projection and
\textbf{exit}
\end{quote}

\textbf{return} the set of projections with the updated best\_jelly as
the optimized\_projection

>>>>

\begin{keyword}
    projection pursuit \sep Jellyfish Search Optimiser (JSO) \sep 
    Optimisation
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

To solve the optimisation problem embedded in the PP guided tour, a
starting projection, an index function and the maximum number of
iterations are provided as input. Then, the current projection is
evaluated by the index function. The projection is then moved in a
direction determined by a random factor, influenced by how far along we
are in the optimization process. Occasionally, completely new directions
may be taken like a jellyfish might with ocean currents. A new
projection is accepted if it is an improvement compared to the current
one, rejected otherwise. This process continues and iteratively improves
the projection, until the pre-specified maximum number of trials is
reached.

\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, breakable, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, coltitle=black, toptitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, leftrule=.75mm, opacityback=0, colback=white, toprule=.15mm, bottomrule=.15mm, rightrule=.15mm, titlerule=0mm]

\textbf{Input}: \texttt{current\_projections}, \texttt{index\_function},
\texttt{trial\_id}, \texttt{max\_trial}

@@

\textbf{Initialize} \texttt{current\_best} as the projection with the
best index value from \texttt{current\_projections}, and
\texttt{current\_idx} as the array of index function values for each
projection in \texttt{current\_projections}

\textbf{for} each trial\_id in 1 to max\_tries \textbf{do}

\begin{quote}
Calculate \(c_t\) based on the current\_idx and max\_trial
\end{quote}

@@

\begin{quote}
Define trend based on the current\_best and current\_projections
\end{quote}

@@

\begin{quote}
Slightly adjust each projection with a small random factor (passive)
\end{quote}

@@

\begin{quote}
For each projection, compare with a random jelly and adjust towards or
away from it (active)
\end{quote}
\end{quote}

@@

\begin{quote}
\textbf{if} any new projection is worse than the current, revert to the
\texttt{current\_projections} for that case

\begin{quote}
Determine the projection with the best index value as the new
current\_best
\end{quote}
\end{quote}

\begin{quote}
\textbf{if} trial\_id \(\geq\) max\_trial, print the last best
projection \textbf{exit}
\end{quote}

\textbf{return} the set of projections with the updated current\_best as
the optimized\_projection
